---
title: "Chapter 1"
author: "Ziteng"
output: html_document
---

Four assumptions for linear model:
<Independence, Constant variance, Normality, Linearity>
With these assumption, our hands are tied sometimes.

GLMs is a method for data violates <Normality>

Goal of this course:
Learn about Generalized Least Squares models and Generalized Linear Models.

```{r include=FALSE}
# Load necessary packages
library(tidyverse)
```

```{r}
# Read in data
df <- read.csv("NystedFarms.csv")

head(df)
```
```{r}
# Figure 2 Transect locations (grey lines) and locations with non-zero counts (blue dots). The size of the blue dots are scaled based on the associated count at each location and are shown along the transects.
ggplot() + 
  geom_point(aes(x=XPos, y=YPos), 
             colour='lightgrey', data=df) +
  geom_point(aes(x=XPos, y=YPos, size=log(Count)), 
             colour="#377eb8", data=subset(df, Count>0)) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank())+
  coord_equal()
```
Check assumptions of linear models on the EIA data

Independence > 
The independence of different counts can be a problem as the animals are moving. And we suspect the densities are correlated along transects.

Normality > 
The residual variance must approach zero when count approaches zero (but we can use Poisson distribution for this).

Constant variance >
The variances may have positive relation with counts.

Linearity >


# 1.1 Research Questions
In line with the programme objectives, we are going to ask some questions of the data:

Does density appear to have changed across phases?
What are the best predictors of density and what do these relationships look like?
Do the animals appear to have redistributed across phases? If so, what does that redistribution look like?



# 1.2 Exploratory Data Analysis
```{r}
# Figure 1.3 Mean counts per unit area (with  95% CIs) across phases
# Compute CIs for each construction phase using normal approx.
alpha <- 0.05
z_alpha_2 <- abs(qnorm(alpha/2)) # z_{alpha/2}
stats <- df %>%
  group_by(Phase) %>%
  summarise(Mean=mean(Count),
            Var=var(Count),
            N=n(),
            CI=z_alpha_2*sqrt(Var/N))

# Plot mean and 95% CIs
ggplot(stats, aes(x=Phase, y=Mean, colour=Phase)) +
  geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI), size=1) +
  geom_point(size=4) +
  geom_text(aes(x=Phase, y=Mean+1.2*CI, label=paste0('N=', N))) +
  ylab("Counts")
```

```{r}
# Figure 1.4 Average counts per unit area (with 95% CIs) across years and months (with phases indicated).
# Compute CIs for density using normal approx.
alpha <- 0.05
z_alpha_2 <- abs(qnorm(alpha/2)) # z_{alpha/2}
stats <- df %>%
  group_by(YearMonth, Phase) %>%
  summarise(Mean=mean(Count/Area),
            Var=var(Count/Area),
            N=n(),
            CI=z_alpha_2*sqrt(Var/N))
```
```{r}
# Plot mean and 95% CIs
ggplot(stats, aes(x=YearMonth, y=Mean, colour=Phase)) +
  geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI), size=1) +
  geom_point(size=2) +
  theme(axis.text.x=element_text(angle=90)) +
  ylab("Abundance per unit area") +
  xlab("Year/Month")
```
From figure 1.3, we see there may be fewer animals in phase 3. However, shorter month coverage might account for the significant difference between phase 3 and phase 1&2, because the density might be changing with month or season. We are not sure what caused the difference. And time alone might not be able to explain the variation.


```{r}
# Figure 1.5: Scatterplots for potential model covariates and estimated densities.
# Libraries
library(gridExtra) # to arrange ggplots in a grid

# Create list of plots of density against other covariates
p <- list()
for (covariate in c("XPos", "YPos", "Depth", "DistCoast"))
{
  p[[covariate]] <- ggplot(df, aes_string(x=covariate, 
                                          y="Count/Area")) + 
    geom_point(alpha=1/2) +
    ylab("Abundance per unit area")    
}

# Arrange ggplots in a grid
grid.arrange(grobs=p)
```

There are outliers.



```{r}
library(fields)
par(mfrow=c(2,2))
quilt.plot(df$XPos[df$Phase=="A"], df$YPos[df$Phase=="A"], df$Count[df$Phase=='A']/df$Area[df$Phase=='A'], pch=20, 
           main="Phase A", xlab="X co-ordinate", ylab="Y co-ordinate", 
           nrow=20, ncol=20, asp=1)
quilt.plot(df$XPos[df$Phase=="B"], df$YPos[df$Phase=="B"], df$Count[df$Phase=='B']/df$Area[df$Phase=='B'], pch=20, 
           main="Phase B", xlab="X co-ordinate", ylab="Y co-ordinate", 
           nrow=20, ncol=20, asp=1)
quilt.plot(df$XPos[df$Phase=="C"], df$YPos[df$Phase=="C"], df$Count[df$Phase=='C']/df$Area[df$Phase=='C'], pch=20, 
           main="Phase C", xlab="X co-ordinate", ylab="Y co-ordinate", 
           nrow=20, ncol=20, asp=1)
par(mfrow=c(1,1))
```












