---
title: "Review_GLMs"
author: "Ziteng"
date: "2024-10-14"
output: html_document
---

# Chapter 1 Introduction

```{r message=FALSE, warning=FALSE}
# Data read in
df <- read.csv("NystedFarms.csv")

# Necessay packages
library(tidyverse)

# Exam data
head(df)
```

```{r}
# Figure for distribution of birds counts by coordinates
ggplot() + 
  geom_point(aes(x=XPos, y=YPos),
             color="lightgrey", data=df) +
  geom_point(aes(x=XPos, y=YPos, size=log(Count)),
             color="#377eb8", data=subset(df, Count>0)) + 
  theme_bw() + coord_equal() + 
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank()) + 
  ggtitle("Figure 1.1")
```

The transect locations are represented by grey dots and locations with non-zero counts are represented by blue dots; size of blue dots are scaled based on the number of counts.

### 1.1 Research Questions

- Does density appear to have changed across phases?
- What are the best predictors of density and what do these relationships look like?
- Do the animals appear to have redistributed across phases? If so, what does that redistribution look like?

### 1.2 Exploratory Data Analysis

```{r}
# Compute CIs for each construction phase using normal approx.
alpha <- 0.5
z_alpha_2 <- abs(qnorm(alpha/2)) # z_{alpha/2}
stats <- df %>%
  group_by(Phase) %>%
  summarise(Mean=mean(Count),
            Var=var(Count),
            N=n(),
            CI=z_alpha_2*sqrt(Var/N))

# Plot mean and 95% CIs
ggplot(stats, aes(x=Phase, y=Mean, color=Phase)) + 
  geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI), linewidth=1) +
  geom_point(size=4) + 
  geom_text(aes(x=Phase, y=Mean+1.2*CI, label=paste0("N=", N))) + 
  ylab("Counts") + ggtitle("Figure 1.2")
```






